---
title: "querychat"
institute: "Fred Hutch Cancer Center"
brand: true
execute: 
  echo: true
format: 
  revealjs:
    incremental: false
---

## {querychat}

- Embed a natural language interface into your shinyapp for filtering
- Add context for LLM to understand your data
- Translates queries to SQL
- **Does not share data with LLM providers**

<https://posit-dev.github.io/querychat/r/>

## Why Should You Care?

- Oftentimes, we can't always anticipate queries
- Lets us bridge the gap
- Gives us a reactive data frame that we can use elsewhere in our app

# Demo

## Setup: Add your API key to the .Renviron file

```{r}
#| eval: false
usethis::edit_r_environ()
```

```bash
# Uncomment the appropriate line and fill in the API key below
# ANTHROPIC_API_KEY=""
# OPENAI_API_KEY=""
```

## Select your model

Use `ellmer::models_anthropic()` or `ellmer::models_openai()` to find a model to use

```{r}
ellmer::models_anthropic()
```

## Plug your model it into your app:

```{r}
#| eval: false
library(palmerpenguins)
library(querychat)
data(penguins)
my_data <- penguins
client <- ellmer::chat_anthropic(model="claude-sonnet-4-5")
qc <- querychat(data_source = penguins, client=client)
qc$app()
```


## Adding Context

- Create a `data_description.md` file in your directory (see `querychat/app_data_description/data_desscription.md)

```{r}
#| eval: false
client <- ellmer::chat_anthropic(model="claude-sonnet-4-5")
qc <- querychat(
    data_source = penguins, 
    client=client,
    data_description = "data_description.md")
qc$app()
```

## Adding Context



## Add a Greeting

```{r}
#| eval: false
client <- ellmer::chat_anthropic(model="claude-sonnet-4-5")
qc <- querychat(
    data_source = penguins, 
    client=client,
    data_description = "data_description.md",
    greeting = "Welcome to exploring the world of penguins!")
qc$app()
```

## Tools: Giving people capabilities

```{r}
#| eval: false
client <- ellmer::chat_anthropic(model="claude-sonnet-4-5")
qc <- querychat(
    data_source = penguins, 
    client=client,
    data_description = "data_description.md",
    greeting = "Welcome to exploring the world of penguins!",
    tools=c("query", "filter"))
qc$app()
```

- [Tools](https://posit-dev.github.io/querychat/r/articles/tools.html)


## Adding visualization to the chat

- Requires the `ggsql` package
- Modify the `tool` argument to include `visualize`:

```{r}
#| eval: false
library(palmerpenguins)
library(querychat)
data(penguins)
my_data <- penguins
client <- ellmer::chat_anthropic(model="claude-sonnet-4-5")
qc <- querychat(
    data_source = penguins, 
    client=client,
    data_description = "data_description.md",
    greeting = "Welcome to exploring the world of penguins!",
    tools=c("query", "filter", "visualize"))
qc$app()
```


# How do we integrate it into dashboards?

## Hooking up into visualizations

The QueryChat object has a few methods that act as reactives that you can use to plug in to visualizations:

```r
# put this in the server function to access the reactive
qc_vals <- qc$server()
```

Then you can add this to your code.

- `qc_vals$df()` - the actual filtered data from QueryChat
- `qc_vals$sql()` - the SQL string
- `qc_vals$title()` - The title of the filter

## Example

```{r}
#| eval: false
# in server function
  output$avg_mass <- renderText({
    avg <- mean(qc_vals$df()$body_mass_g, na.rm = TRUE)
    paste0(round(avg, 0), "g")
  })

# in ui function
   value_box(
      title = "Avg Body Mass",
      value = textOutput("avg_mass"),
      showcase = bs_icon("speedometer"),
      theme = "info"
    )
```

```r
runApp("querychat/app_big_example")
```
